---
title: "EDA"
author: "Sam Lee"
format: pdf
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
library(invgamma)
library(ggplot2)
library(gridExtra)

set.seed(12142023)

data_dir = "march-machine-learning-mania-2023/"
```

```{r message=F}
season_results = read_csv(str_c(data_dir, "MRegularSeasonDetailedResults.csv"))
teams = read_csv(str_c(data_dir, "MTeams.csv"))
tourney_results = read_csv("2023 Game Data.csv")

transform_data = function(t){
  t %>% pivot_longer(cols=c(WTeamID, LTeamID), names_to="WL", values_to = "TeamID") %>%
  select(TeamID, WL, WScore, LScore, WFGM, WFGA, LFGM, LFGA) %>%
  left_join(teams %>% select(TeamID, TeamName)) %>%
  pivot_longer(cols=c(WScore, LScore), values_to = "Score", names_to = "WLScore") %>%
  pivot_longer(cols=c(WFGM, LFGM), values_to="FGM", names_to = "WLFGM") %>%
  pivot_longer(cols=c(WFGA, LFGA), values_to="FGA", names_to = "WLFGA") %>%
  mutate(WL = sapply(WL, function(x)substr(x,1,1)),
         WLScore = sapply(WLScore, function(x)substr(x,1,1)),
         WLFGM = sapply(WLFGM, function(x)substr(x,1,1)),
         WLFGA = sapply(WLFGA, function(x)substr(x,1,1))
  ) %>% rowwise() %>%
  filter(all(c(WL, WLScore, WLFGM, WLFGA) == 
               first(c(WL, WLScore, WLFGM, WLFGA)))) %>%
  select(TeamID, WL, TeamName, Score, FGM, FGA)
}

#2022 Season
season.2022 <- season_results %>% 
  filter(Season %in% 2022)

season.2022 <- transform_data(season.2022)
  
season.2022 %>% mutate(
    FGP = FGM/FGA
  ) -> season.2022

calculate_prior_fgp = function(p, beta=1){
  #Returns alpha for a Beta(alpha, beta) such that alpha/(alpha+beta) = p (expected value)
  return(p*beta/(1-p))
}


#Calculate priors for the field goal percentage ~ Beta(alpha, beta)
#and for the field goal attempts ~ N(mu, sigma^2)
season.2022 %>% group_by(TeamID) %>%
  summarize(
    fgp.alpha.prior = calculate_prior_fgp(mean(FGP)),
    fgp.beta.prior = 1,
    fga.lambda.prior = mean(FGA),
    fga.tau.prior = (max(FGA)-min(FGA))/3,
    #Method of Moments https://arxiv.org/pdf/1605.01019.pdf
    fga.gamma.prior = mean(FGA)^2/var(FGA)+2,
    fga.phi.prior = mean(FGA)*(mean(FGA)^2/var(FGA)+1)
  ) -> season.2022.priors

#2023 Season we want to model
season.2023 <- season_results %>% 
  filter(Season %in% 2023)

season.2023 <- transform_data(season.2023)

#Calculate posteriors for FGP for the 2023 season
season.2023.posteriors = season.2023 %>% left_join(season.2022.priors, by=join_by(TeamID)) %>%
  group_by(TeamID) %>%
  summarize (
    fgp.alpha.posterior = sum(FGM)+first(fgp.alpha.prior),
    fgp.beta.posterior = sum(FGA)-sum(FGM)+first(fgp.beta.prior),
    fga.lambda.prior = first(fga.lambda.prior),
    fga.tau.prior = first(fga.tau.prior),
    fga.gamma.prior = first(fga.gamma.prior),
    fga.phi.prior = first(fga.phi.prior)
  )

#Gibbs Sampling Method to Define Posterior
posterior.matrix = as.matrix(season.2023.posteriors[c("fga.lambda.prior", "fga.tau.prior",                                          
                                                      "fga.gamma.prior", "fga.phi.prior")])

iterations = 10000

#Matrices to store posterior distributions
posterior.normal.matrix = matrix(ncol=iterations, nrow=nrow(posterior.matrix))
posterior.invgamma.matrix = matrix(ncol=iterations, nrow=nrow(posterior.matrix))

#Calculate the Normal posterior distribution for each ith team via Gibbs sampling
for(i in 1:nrow(posterior.matrix)){
  ith_team = posterior.matrix[i,]
  data_i = season.2023[season.2023$TeamID == as.numeric(season.2023.posteriors[i,"TeamID"]), "FGA"] %>%
    unlist()
  
  #Gibbs sampling algorithm
  burn = 100
  iters <- iterations + burn
  mu.save <- rep(0, iters)
  mu.save <- ith_team["fga.lambda.prior"]
  sigma2.save <- rep(0, iters)
  sigma2 = ith_team["fga.phi.prior"]/(ith_team["fga.gamma.prior"]-1)
  sigma2.save[1] = sigma2
  
  lambda = ith_team["fga.lambda.prior"]
  tau = ith_team["fga.tau.prior"]
  gamma = ith_team["fga.gamma.prior"]
  phi = ith_team["fga.phi.prior"]
  n = length(data_i)
  
  if(any(is.na(ith_team))){
    posterior.normal.matrix[i,] = rep(NA_real_, iterations)
    posterior.invgamma.matrix[i,] = rep(NA_real_, iterations)
  } else {
    for(t in 2:iters){
      #Full conditional of mu
      lambda.p <- (tau^2*sum(data_i) + sigma2*lambda)/(tau^2*n + sigma2)
      tau2.p <- sigma2*tau^2/(tau^2*n + sigma2)
      
      #New value of mu
      mu <- rnorm(1, lambda.p, sqrt(tau2.p))
      mu.save[t] <- mu
      
      #Full conditional of sigma2
      gamma.p <- gamma + length(data)/2
      phi.p <- phi + sum((data_i - mu)^2)/2
      
      #New value of sigma2
      sigma2 <- rinvgamma(1, gamma.p, phi.p)
      sigma2.save[t] <- sigma2
    }
    
    posterior.normal.matrix[i,] = mu.save[-(1:burn)]
    posterior.invgamma.matrix[i,] = sigma2.save[-(1:burn)]
  }
  
  #print(i)
  
}


season.2023.posteriors$fga.mu.posterior = rowMeans(posterior.normal.matrix)
season.2023.posteriors$fga.sigma.posterior = sqrt(rowMeans(posterior.invgamma.matrix))

season.2023.posteriors %>% 
  filter(!is.na(fga.mu.posterior)) -> season.2023.posteriors

#Monte Carlo Simulation to Simulate FGM
posterior.fgm.matrix = matrix(ncol=iterations, nrow=nrow(season.2023.posteriors))
for(i in 1:nrow(season.2023.posteriors)){
  #Randomly sample from p from the posterior beta distribution on Field Goal Percentage
  p = rbeta(iterations, as.numeric(season.2023.posteriors[i, "fgp.alpha.posterior"]), 
            as.numeric(season.2023.posteriors[i, "fgp.beta.posterior"]))
  #Calculate distribution of mean FGM by multiplying p by a random sample of FGA by team i
  f = rnorm(iterations, as.numeric(season.2023.posteriors[i, "fga.mu.posterior"]), 
            as.numeric(season.2023.posteriors[i, "fga.sigma.posterior"]))
  
  posterior.fgm.matrix[i,] = p*f
}

tourney_results[c("SEED", "TEAM...3")] %>% 
  setNames(c("Seed", "TeamName")) -> tourney_results
clean_team_names = function(t){
  t$TeamName = sapply(t$TeamName, function(x){
    x = x %>% str_replace("[.]", "")
    x = x %>% str_replace("Florida", "FL")
    if(x == "Saint Mary's")x = "St Mary's CA"
    if(x == "College of Charleston")x = "Col Charleston"
    if(x == "Louisiana Lafayette")x = "Lafayette"
    if(x == "Fairleigh Dickinson")x = "F Dickinson"
    if(x == "Northern Kentucky")x = "N Kentucky"
    if(x == "Southeast Missouri St")x = "SE Missouri St"
    if(x == "Texas A&M Corpus Chris")x = "TAM C. Christi"
    if(x == "Texas Southern")x = "TX Southern"
    if(x == "Montana St")x = "Montana St"
    if(x == "Kennesaw St")x = "Kennesaw"
    if(x == "Kent St")x = "Kent"
    if(x == "North Carolina St")x = "NC State"
    return(x)
  })
  return(t)
}
tourney_results = clean_team_names(tourney_results)

tourney_results %>% 
  left_join(teams[c("TeamID", "TeamName")]) -> tourney_results

#Omit the first four
tourney_results %>% 
  filter(!TeamName %in% c("TX Southern", "Nevada",
                          "Mississippi St", "SE Missouri St")) %>% 
  distinct() -> tourney_results

tourney_results$Region = rep(c("E", "S", "W", "M"), each=2, times=8)
```

```{r}
#2023 Tournament Simulation

regions = c("E", "S", "W", "M")
tourney_results %>% group_by(Region) %>%
  mutate(
    Order = rep(LETTERS[1:(n()/2)], each=2)
  ) -> tourney_results

matchups = tibble()

compare_teams = function(k, l, alpha=0.25){
  k = which(season.2023.posteriors$TeamID == k)
  l = which(season.2023.posteriors$TeamID == l)
  list(
    p = mean(posterior.fgm.matrix[k,] > posterior.fgm.matrix[l,]),
    q = quantile(posterior.fgm.matrix[k,] - posterior.fgm.matrix[l,], alpha)
  )
}

tourney_results$Round = 1

for(round in 1:4){
 for(region in regions){
    t = tourney_results
    if(round > 1)t = matchups
    
    if(round < 5){
      #These are all the regional matches
      region.subset = t %>% 
        filter(Region == region & Round == round) 
    }
    
    region.subset$p = NA_real_
    region.subset$alpha.probability = NA_real_
    region.subset$Round = round+1
    if(round > 1){
      half = region.subset$Order[1:(length(region.subset$Order)/2)]
      region.subset$Order = c(half, rev(half))
      
      matchups[matchups$Region == region & matchups$Round == round,
               ]$Order =c(half, rev(half))
    }
    
    
    region.subset %>% 
      arrange(Order) -> region.subset
    
    #Loop through every game
    i = 1
    while(i < nrow(region.subset)){
      p = compare_teams(region.subset[i,]$TeamID,
                        region.subset[i+1,]$TeamID)[["p"]]
      #Predictive probability distribution is a Bernoulli Distribution
      if(p > (1-p)){
        region.subset[i,]$p = p
        region.subset[i,]$alpha.probability = 
          compare_teams(region.subset[i,]$TeamID,
                region.subset[i+1,]$TeamID)[["q"]] %>% as.vector() > 0
        matchups = rbind(matchups, region.subset[i,])
      } else {
        region.subset[i+1,]$p = 1-p
        region.subset[i+1,]$alpha.probability = 
          compare_teams(region.subset[i+1,]$TeamID,
                region.subset[i,]$TeamID)[["q"]] %>% as.vector() > 0
        matchups = rbind(matchups, region.subset[i+1,])
      }
      i = i + 2
    }
  }
}

#Final Four and Champtionship
for(round in 5:6){
  t.subset = matchups %>% 
    filter(Round == round)
  t.subset$Round = round+1
  
  #Loop through every game
  i = 1
  while(i < nrow(t.subset)){
    p = compare_teams(t.subset[i,]$TeamID,
                      t.subset[i+1,]$TeamID)[["p"]]
    #Predictive probability distribution is a Bernoulli Distribution
    if(p > (1-p)){
      t.subset[i,]$p = p
      t.subset[i,]$alpha.probability = compare_teams(t.subset[i,]$TeamID,
              t.subset[i+1,]$TeamID)[["q"]] %>% as.vector() > 0      
      matchups = rbind(matchups, t.subset[i,])
    } else {
      t.subset[i+1,]$p = 1-p
      t.subset[i+1,]$alpha.probability = compare_teams(t.subset[i+1,]$TeamID,
              t.subset[i,]$TeamID)[["q"]] %>% as.vector() > 0
      matchups = rbind(matchups, t.subset[i+1,])
    }
    i = i + 2
  }
}

```


## 2023 Tournament Simulation

The column *p* indicates the predictive posterior probability of how likely that team was to make more field goals than their opposing team in the previous round.

#### First Round Matchups

```{r echo=F}
tourney_results[c("Seed", "TeamName", "Region")] %>% knitr::kable()
```

#### Second Round Matchups

```{r echo=F}
matchups %>%
  select(Seed, TeamName, Region, p, Round, Order) %>%
  arrange(Round, Order) %>%
  select(-Order) -> matchups

matchups %>% 
  filter(Round == 2) %>% 
  select(-Round) %>%
  knitr::kable()
```

#### Sweet 16

```{r echo=F}
matchups %>% 
  filter(Round == 3) %>% 
  select(-Round) %>%
  knitr::kable()
```

#### Elite 8

```{r echo=F}
matchups %>% 
  filter(Round == 4) %>% 
  select(-Round) %>%
  knitr::kable()
```

#### Final Four

```{r echo=F}
matchups %>% 
  filter(Round == 5) %>% 
  select(-Round) %>%
  knitr::kable()
```

#### Championship

```{r echo=F}
matchups %>% 
  filter(Round == 6) %>% 
  select(-Round) %>%
  knitr::kable()
```

#### Champion

```{r echo=F}
matchups %>% 
  filter(Round == 7) %>% 
  select(-Round) %>%
  knitr::kable()
```

Prior Distributions for Oral Roberts and Gonzaga

#### Field Goal Percentage

$\text{FGP}_k \sim \text{Beta}(\alpha_k, \beta_k) \quad \forall k \in \text{Teams}$

```{r}
k = 1211 #Gonzaga
l = 1331 #Oral Robers

k.alpha = season.2022.priors %>% 
  filter(TeamID == k) %>% pull(fgp.alpha.prior)
k.beta = season.2022.priors %>% 
  filter(TeamID == k) %>% pull(fgp.beta.prior)

l.alpha = season.2022.priors %>% 
  filter(TeamID == l) %>% pull(fgp.alpha.prior)
l.beta = season.2022.priors %>% 
  filter(TeamID == l) %>% pull(fgp.beta.prior)

ggplot(data = data.frame(x = c(0, 1)), aes(x)) +
  stat_function(fun = dbeta, n = 1001, 
                args = list(shape1 = k.alpha, shape2 = k.beta),
                aes(color = "Gonzaga"), 
                show.legend=T) +
  stat_function(fun = dbeta, n = 1001, 
                args = list(shape1 = l.alpha, shape2 =l.beta), 
                aes(color = "Oral Roberts"), show.legend=T) +
  ylab(expression(pi(FGP[k]))) +
  xlab(expression(FGP[k])) +
  ggtitle("Prior Distributions") +
  theme_minimal() +
  labs(color = "Distribution") +
  scale_color_manual(
    values = c("Gonzaga" = "steelblue", "Oral Roberts" = "coral")) +
  theme(legend.position = "top")
```

#### Field Goal Attempts

$\text{FGA}_{k} \sim N(\lambda_{k}, \tau^2_{k})$

```{r}
k = 1211 #Gonzaga
l = 1331 #Oral Robers

k.lambda = season.2022.priors %>% 
  filter(TeamID == k) %>% pull(fga.lambda.prior)
k.tau = season.2022.priors %>% 
  filter(TeamID == k) %>% pull(fga.tau.prior)

l.lambda = season.2022.priors %>% 
  filter(TeamID == l) %>% pull(fga.lambda.prior)
l.tau = season.2022.priors %>% 
  filter(TeamID == l) %>% pull(fga.tau.prior)

ggplot(data = data.frame(x = c(0, 100)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, 
                args = list(mean = k.lambda, sd = k.tau),
                aes(color = "Gonzaga"), 
                show.legend=T) +
  stat_function(fun = dnorm, n = 1001, 
                args = list(mean = l.lambda, sd =l.tau), 
                aes(color = "Oral Roberts"), show.legend=T) +
  ylab(expression(pi(FGA[k]))) +
  xlab(expression(FGA[k])) +
  ggtitle("Prior Distributions") +
  theme_minimal() +
  labs(color = "Distribution") +
  scale_color_manual(
    values = c("Gonzaga" = "steelblue", "Oral Roberts" = "coral")) +
  theme(legend.position = "top")
```

Posterior Distributions for Oral Roberts and Gonzaga

#### Field Goal Percentage

$\text{FGP}_k|\text{Data}_{k} \sim \text{Beta}(\alpha_k, \beta_k) \quad \forall k \in \text{Teams}$

```{r}
k = 1211 #Gonzaga
l = 1331 #Oral Robers

k.alpha = season.2023.posteriors %>% 
  filter(TeamID == k) %>% pull(fgp.alpha.posterior)
k.beta = season.2023.posteriors %>% 
  filter(TeamID == k) %>% pull(fgp.beta.posterior)

l.alpha = season.2023.posteriors %>% 
  filter(TeamID == l) %>% pull(fgp.alpha.posterior)
l.beta = season.2023.posteriors %>% 
  filter(TeamID == l) %>% pull(fgp.beta.posterior)

ggplot(data = data.frame(x = c(0.25, 0.75)), aes(x)) +
  stat_function(fun = dbeta, n = 1001, 
                args = list(shape1 = k.alpha, shape2 = k.beta),
                aes(color = "Gonzaga"), 
                show.legend=T) +
  stat_function(fun = dbeta, n = 1001, 
                args = list(shape1 = l.alpha, shape2 =l.beta), 
                aes(color = "Oral Roberts"), show.legend=T) +
  ylab(expression(pi(FGP[k]))) +
  xlab(expression(FGP[k])) +
  ggtitle("Posterior Distributions") +
  theme_minimal() +
  labs(color = "Distribution") +
  scale_color_manual(
    values = c("Gonzaga" = "steelblue", "Oral Roberts" = "coral")) +
  theme(legend.position = "top")
```

We estimated the following the posterior distributions for their Field Goal Percentage:

$\text{FGP}_\text{Gonzaga}|\text{Data}_\text{Gonzaga} \sim \text{Beta}(1026.116, 939)$

$\text{FGP}_\text{Oral Roberts}|\text{Data}_\text{Oral Roberts} \sim \text{Beta}(896.829, 988)$

Hence,

$E(\text{FGP}_\text{Gonzaga}|\text{Data}_\text{Gonzaga})=0.522$

$E(\text{FGP}_\text{Oral Roberts}|\text{Data}_\text{Oral Roberts})=0.476$

$V(\text{FGP}_\text{Gonzaga}|\text{Data}_\text{Gonzaga})$=`r k.alpha*k.beta/((k.alpha+k.beta+1)*(k.alpha+k.beta)^2)`

$V(\text{FGP}_\text{Oral Roberts}|\text{Data}_\text{Oral Roberts})$=`r l.alpha*l.beta/((l.alpha+l.beta+1)*(l.alpha+l.beta)^2)`

#### Field Goal Attempts

$\text{FGA}_{k}|\text{Data}_k \sim N(\mu_{k}, \sigma^2_{k})$`

```{r}
k = 1211 #Gonzaga
l = 1331 #Oral Robers

k = which(season.2023.posteriors$TeamID == k)
l = which(season.2023.posteriors$TeamID == l)

k.mu = mean(posterior.normal.matrix[k,])
k.sigma = sqrt(mean(posterior.invgamma.matrix[k,]))

l.mu = mean(posterior.normal.matrix[l,])
l.sigma = sqrt(mean(posterior.invgamma.matrix[l,]))

ggplot(data = data.frame(x = c(0, 100)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, 
                args = list(mean = k.mu, sd = k.sigma),
                aes(color = "Gonzaga"), 
                show.legend=T) +
  stat_function(fun = dnorm, n = 1001, 
                args = list(mean = l.mu, sd =l.sigma), 
                aes(color = "Oral Roberts"), show.legend=T) +
  ylab(expression(pi(FGA[k]))) +
  xlab(expression(FGA[k])) +
  ggtitle("Posterior Distributions") +
  theme_minimal() +
  labs(color = "Distribution") +
  scale_color_manual(
    values = c("Gonzaga" = "steelblue", "Oral Roberts" = "coral")) +
  theme(legend.position = "top")
```

We estimated the following for the posterior distributions for Gonzaga and Oral Roberts for their Field Goal Attempts:

$\text{FGA}_\text{Gonzaga}|\text{Data}_{\text{Gonzaga}} \sim N(61.37, 8.45^2)$

$\text{FGA}_\text{Oral Roberts}|\text{Data}_{\text{Oral Roberts}} \sim N(62.86, 8.44^2)$

#### Posterior Predictive Distribution on FGW

$\text{FGM}_{k}|\text{Data}_k > \text{FGM}_l|\text{Data}_k \sim \text{Bernoulli}(p_{kl})$

```{r}
k = 1211 #Gonzaga
l = 1331 #Oral Robers

k = which(season.2023.posteriors$TeamID == k)
l = which(season.2023.posteriors$TeamID == l)

p.kl = data.frame(Event = c("Predicted Win", "Predicted Loss"), 
                  Probability = c(mean(posterior.fgm.matrix[k,] > posterior.fgm.matrix[l,]), 
                                                         1 - mean(posterior.fgm.matrix[k,] > posterior.fgm.matrix[l,])))

ggplot(p.kl, aes(x = Event, y = Probability, fill = Event)) +
  geom_bar(stat = "identity") +
  labs(title = "Predictive Posterior Bernoulli Distribution", x = "FGM of Gonzaga > FGM of Oral Roberts", y = "Probability") +
  scale_fill_manual(values = c("Predicted Win" = "steelblue", "Predicted Loss" = "coral"))+
  theme_minimal()+
  theme(legend.position = "none")

plot(density(posterior.fgm.matrix[k,] - posterior.fgm.matrix[l,]), main="Posterior Distribution of Difference in FGM")
```

If we assume that the FGA is independent of FGP, then,

$E(\text{FGM}_\text{Gonzaga}|\text{Data}_\text{Gonzaga})=E(\text{FGP}_\text{Gonzaga}|\text{Data}_\text{Gonzaga})E(\text{FGA}_\text{Gonzaga}|\text{Data}_\text{Gonzaga})$ =`r (k.alpha/(k.alpha+k.beta))*k.mu`

$E(\text{FGM}_\text{Oral Roberts}|\text{Data}_\text{Oral Roberts})=E(\text{FGP}_\text{Oral Roberts}|\text{Data}_\text{Oral Roberts})E(\text{FGA}_\text{Oral Roberts}|\text{Data}_\text{Roberts})$ =`r (l.alpha/(l.alpha+l.beta))*l.mu`

Hence, 

$E(\text{FGM}_\text{Gonzaga}|\text{Data}_\text{Gonzaga}-\text{FGM}_\text{Oral Roberts}|\text{Data}_\text{Oral Roberts})$=$E(\text{FGM}_\text{Gonzaga}|\text{Data}_\text{Gonzaga})-E(\text{FGM}_\text{Oral Roberts}|\text{Data}_\text{Oral Roberts})$=`r (k.alpha/(k.alpha+k.beta))*k.mu-(l.alpha/(l.alpha+l.beta))*l.mu`

Where the expected values from FGA were approximated using Gibbs sampling techniques.

Our estimated posterior predictive mean using Monte Carlo Approximation is `r mean(posterior.fgm.matrix[k,] - posterior.fgm.matrix[l,])`.

Our estimated posterior predictive variance using Monte Carlo Approximation is `r var(posterior.fgm.matrix[k,] - posterior.fgm.matrix[l,])`